\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2018-11-14 draft.tex
polish

MapReduce is one of the most popular framework for data parallel computation.
A MapReduce~\cite{dean04,hadoop} program consists of several pairs of \emph{mappers} and \emph{reducers} running on a machine cluster for handling big data in parallel. Usually mappers and reducers are the only components in a MapReduce program that involves concurrency. Mappers read data from a distributed database and output a sequence of \emph{key-value} pairs. The sequence elements (key-value pairs) with the same key will be sent to the same reducer for further processing. Due to scheduling polices and network latency, the same inputs may arrive at a reducer in different order in different executions, which results in the so-called \emph{commutativity problem} of reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}, that is, if the output of a reducer is independent of the order of its inputs.
If all reducers are commutative, they will have the same external behavior under all possible schedules. In such a case, by fixing a schedule, the verification problem of a MapReduce program can be reduced to the verification problem of a sequential program, which is known to be much easier then the verification problem of concurrent programs. One needs to consider only a particular interleaving of input values instead of all of them.

@@

In this paper, we propose heuristics for the commutativity problem that works very efficiently on a large set of practical integer reducer programs. We reduce the commutativity problem to an SMT problem in a sound but incomplete manner. We also propose optimizations to enable the approach working on real world examples. For the case that the reducer is not proven commutative, we complement the approach by using testing to find concrete counterexamples. We implement all these heuristics as a tool called J-ReCoVer (Java Reducer Commutativity Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|.

We use two sets of integer reducer benchmarks to evaluate J-ReCoVer. The first set is collected from open repositories such as Github and Bitbuckect. With the help of a search engine~\cite{searchcode} over those repositories, we collected 118 examples. J-ReCoVer is able to correctly analyze all but 3 of them. We also tried all 10 integer reducers descried in verification literature~\cite{ChenHSW15,ChenSW16} and J-ReCoVer can handle all of them.

\subsection*{Related Works}
Reducer commutativity problem can be reduced to program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There is a series of research results on program equivalence checking (or closely related topics such as regression verification and translation validation)~\cite{Pnueli:1998:TV,symdiff,fedyukovich2015automated,sharma2013data,godlin2009regression,fedyukovich2016property,felsing2014automating,KieferKlebanovUlbrich2017,lahiri2013differential,grossman2017verifying,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of research effort in this direction is in finding good synchronization points. For the case of commutativity analysis, one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

One main motivation of commutativity analysis is to reduce the verification problem of a concurrent program to a sequential one. There are several researches sharing the similar motivation, e.g., the concept of \emph{robustness} of event-driven asynchronous programs~\cite{ahmed2017:robustness}, \emph{robustness} of program running under relaxed memory model~\cite{ahmed2013:robustness,AbdullaACLR13,AbdullaACLR12}, and \emph{synchronizability} of communicating finite state machines~\cite{FinkelL17,basu2012synchronizability,basu2011choreography}. However, their computation models are quite different to that of reducers and hence the result cannot be directly applied. Besides verification, another interesting research direction is the synthesis of MapReduce programs~\cite{SmithA16}. Commutativity analysis is used as a component under the synthesis framework.

@@

Here $v\in \Var$ are variables and $c \in \Z$ are constant integer values, respectively. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{guard} in $\Grd$ can be a comparison between two expressions or boolean combinations of several guards. A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. A few examples of reducer programs in this model can be found in Figure~\ref{fig:reducer_example}.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot~\cite{soot} to further convert it to the so called ``Jimple'' format, which is an intermediate language that is designed to simplify the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component so that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some program transformation tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to an special assignment (will be explained in Section~\ref{sec:program_trans1} and~\ref{sec:program_trans2}).

\textbf{The BugFinder component} randomly generated 1500 pairs of lists. The lists in a pair are permutations of each other. A concrete counterexample is reported if inputting any pair of lists from the 1500 generated ones to the reducer generates different results.
Our procedure for generating random pairs is quite naive. We use five different input list of lengths 5, 7, 9 ,11, and 13. For each length, we generate 300 lists and pick uniformly at random one of its permutations. Although the approach is simple, in practice it finds counterexamples in all of our non-commutative benchmarks.

\textbf{The Prover component} reduces the commutativity problem to an SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to an SMT problem and give it to the SMT solver Z3~\cite{z3} for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative.
However, if we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enables the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate SMT formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

The validity of Equation~(\ref{eq:commu}) implies that swapping any two consecutive input values does not change the final variable valuation. Informally, it says that starting from the same initial valuations of variables and with two different input orders, $[n_1;n_2]$ and $[n_2;n_1]$, the values of all program variables are the same after we execute the loop body twice. The first execution reads $n_1$ and then reads $n_2$. The other execution reads the two inputs in the reverse order.

>>>>

MapReduce is one of the most popular frameworks for data parallel computation.
A MapReduce~\cite{dean04,hadoop} program consists of several pairs of \emph{mappers} and \emph{reducers} running on a machine cluster for handling big data in parallel. Usually, mappers and reducers are the only components in a MapReduce program that involves concurrency. Mappers read data from a distributed database and output a sequence of \emph{key-value} pairs. The sequence elements (key-value pairs) with the same key will be sent to the same reducer for further processing. Due to scheduling policies and network latency, the same inputs may arrive at a reducer in a different order in different executions, which results in the so-called \emph{commutativity problem} of reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}, that is, if the output of a reducer is independent of the order of its inputs.
If all reducers are commutative, they will have the same external behavior under all possible schedules. In such a case, by fixing a schedule, the verification problem of a MapReduce program can be reduced to the verification problem of a sequential program, which is known to be much easier than the verification problem of concurrent programs. One needs to consider only a particular interleaving of input values instead of all of them.

@@

In this paper, we propose heuristics for the commutativity problem that works very efficiently on a large set of practical integer reducer programs. We reduce the commutativity problem to an SMT problem in a sound but incomplete manner. We also propose optimizations to enable the approach that works on real-world examples. For the case that the reducer is not proven commutative, we complement the approach by using testing to find concrete counterexamples. We implement all these heuristics as a tool called J-ReCoVer (Java Reducer Commutativity Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|.

We use two sets of integer reducer benchmarks to evaluate J-ReCoVer. The first set is collected from open repositories such as Github and Bitbucket. With the help of a search engine~\cite{searchcode} over those repositories, we collected 118 examples. J-ReCoVer is able to correctly analyze all but 3 of them. We also tried all 10 integer reducers described in verification literature~\cite{ChenHSW15,ChenSW16} and J-ReCoVer can handle all of them.

\subsection*{Related Works}
Reducer commutativity problem can be reduced to a program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There is a series of research results on program equivalence checking (or closely related topics such as regression verification and translation validation)~\cite{Pnueli:1998:TV,symdiff,fedyukovich2015automated,sharma2013data,godlin2009regression,fedyukovich2016property,felsing2014automating,KieferKlebanovUlbrich2017,lahiri2013differential,grossman2017verifying,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of research effort in this direction is in finding good synchronization points. For the case of commutativity analysis, one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

One main motivation of commutativity analysis is to reduce the verification problem of a concurrent program to a sequential one. There are several pieces of research sharing the similar motivation, e.g., the concept of \emph{robustness} of event-driven asynchronous programs~\cite{ahmed2017:robustness}, the \emph{robustness} of program running under relaxed memory model~\cite{ahmed2013:robustness,AbdullaACLR13,AbdullaACLR12}, and \emph{synchronizability} of communicating finite state machines~\cite{FinkelL17,basu2012synchronizability,basu2011choreography}. However, their computation models are quite different from that of reducers and hence the result cannot be directly applied. Besides verification, another interesting research direction is the synthesis of MapReduce programs~\cite{SmithA16}. Commutativity analysis is used as a component under the synthesis framework.

@@

Here any $v\in \Var$ is a variable and any $c \in \Z$ is a constant integer value. An \emph{integer expression} in $\Exp$ can be either a variable, constant value, a call to the $\cur$ function that reads and consumes an input value of the reducer, their combinations over basic arithmetic operations, or a non-deterministically chosen integer value $*$.
A \emph{guard} in $\Grd$ can be a comparison between two expressions or boolean combinations of several guards. A \emph{command} in $\Cmd$ can be an assignment, a branch statement, concatenation of a sequence of commands, or $out(v)$ that outputs the value of $v$.
A \emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$, where $s_1,s_2,s_3 \in \Cmd$, and $s_2$ is assumed to have at most one $\cur$ expression in each \emph{branch}\footnote{We define a \emph{branch} of a command $s$ as a sequence obtained by replacing every branch statement $\ite{g}{s_1}{s_2}$ in $s$ with either $(g); s_1$ or $(not\ g) ;s_2$. For example, the loop body of Figure~\ref{fig:reducer_example}(a) has two branches $t:=\cur; (t>0); m:=t$ and $t:=\cur;(not\ t>0)$ }. According to our observation over hundreds  of reducer programs in open repositories, reducer programs are almost always in this form. The $\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until the entire input list is consumed. A few examples of reducer programs in this model can be found in Figure~\ref{fig:reducer_example}.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot~\cite{soot} to further convert it to the so-called ``Jimple'' format, which is an intermediate language that is designed to simplify the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component so that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some program transformation tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to a special assignment (will be explained in Section~\ref{sec:program_trans1} and~\ref{sec:program_trans2}).

\textbf{The BugFinder component} randomly generated 1500 pairs of lists. The lists in a pair are permutations of each other. A concrete counterexample is reported if inputting any pair of lists from the 1500 generated ones to the reducer generates different results.
Our procedure for generating random pairs is quite naive. We use five different input list of lengths 5, 7, 9, 11, and 13. For each length, we generate 300 lists and pick uniformly at random one of its permutations. Although the approach is simple, in practice it finds counterexamples in all of our non-commutative benchmarks.

\textbf{The Prover component} reduces the commutativity problem to an SMT formula satisfiability problem. At a high level, we check equivalence between a reducer program and its variant that has two consecutive inputs swapped. We show that this can be reduced to an SMT problem and give it to the SMT solver Z3~\cite{z3} for solving. In case that Z3 proved the formula unsatisfiable, we know that swapping any two consecutive inputs of the reducer will not change its output. Because all permutations of a list can be obtained by swapping consecutive elements finitely many times, it follows that the reducer outputs the same value for all permutations of the same list of inputs.
In this case,  J-ReCoVer stops and reports that the reducer is commutative.
However, if we run the approach in a naive manner, it does not scale to large programs and is very imprecise.  We will introduce two critical optimizations that enable the proposed approach working for real-world examples in Section~\ref{section:optimizations}: a data flow analysis to extract critical variables and an improved symbolic execution algorithm to generate SMT formulae. More details of the algorithms used in the Preprocessor and the Prover components will be explained in Section~\ref{sec:preprocessor_prover} and the optimizations will be introduced in~\ref{section:optimizations}.

@@

The validity of Equation~(\ref{eq:commu}) implies that swapping any two consecutive input values will not change the final variable valuation. Informally, it says that starting from the same initial valuations of variables and with two different input orders, $[n_1;n_2]$ and $[n_2;n_1]$, the values of all program variables are the same after we execute the loop body twice. The first execution reads $n_1$ and then reads $n_2$. The other execution reads the two inputs in the reverse order.
